---
title: "soil cores 2021 nwere"
author: "Becca Evans"
date: '2022-04-15'
output: html_document
---



```{r setup, include=FALSE}
setwd("C:/Users/becca/Dropbox/Becca-John PhD share/DATA/2021 soils")
#packages
library(tidyverse)
library(dplyr)
library(lme4)
library(emmeans)#constrasts
library(lmerTest)
library(ggplot2)
library(MASS)
library(car) #multicollinarity 
```

```{r}
#data prep

CN <- read.csv("nwere_2018_2021soils_allparameters.csv")

#remove 8 
CN_all <- CN %>% filter(n.treat!="8")
head(CN_all)
names(CN_all)
names(CN_all)[names(CN_all) == "X21_d13c"] <- "d13c_21"
names(CN_all)[names(CN_all) ==  "X21_mg.c.gsoil"] <- "C_21"
names(CN_all)[names(CN_all) ==  "X21_d15n"] <- "d15n_21"
names(CN_all)[names(CN_all) ==  "X21_mg.n.gsoil"] <- "N_21"
names(CN_all)[names(CN_all) ==  "X18_mg.c.gsoil"] <- "C_18"
names(CN_all)[names(CN_all) ==  "X18_mg.n.gsoil"] <- "N_18"




CN_all$replicate <- factor(CN_all$replicate)
CN_all$n.treat <- factor(CN_all$n.treat)
CN_all$plot <- factor(CN_all$plot)
CN_all$block <- factor(CN_all$block)
str(CN_all)
names(CN_all)

```

means C,N d13c, d15n no rep for figs include BD becuase this was the only thing sig use all data
```{r}
names(CN)
std.error <- function(x) sqrt(var(x,na.rm=TRUE)/length(na.omit(x)))
soil.summary <- CN %>% group_by(n.treat, weevil, site,block)%>%
  summarise_at(c("X21_d15n","X21_mg.n.gsoil","X18_mg.c.gsoil","X18_mg.n.gsoil","X21_d13c","X21_mg.c.gsoil", "bd.pits.g.cm3.2018","sum.13","sum.14","sum.21","litter_kg_m2_mean","herb_kg_m2_mean","som_kg_m2_mean","litter_kg_m2_std.error" ,"herb_kg_m2_std.error" ,"som_kg_m2_std.error","numberwillow","numberother","totalnumber","willowstemdiam","otherstemdiam","totalstemdiam"), funs(mean, std.error))

dodge <- position_dodge(.3)
```
 
 Select only primary 
 
 #clean data
 
```{r}
CN$replicate <- factor(CN$replicate)
CN$n.treat <- factor(CN$n.treat)
CN$plot <- factor(CN$plot)
CN$block <- factor(CN$block)
str(CN)

```
 

```{r}
pri_cn_rep <- filter(CN,site=="East"|site=="West" )
pri_cn_norep <- filter(soil.summary, site=="East"|site=="West")
```

Select only secondary

```{r}
sec_cn_rep <- filter(CN,site=="Norway"|site=="Harmony" )
sec_cn_norep <- filter(soil.summary, site=="Norway"|site=="Harmony")
```


Figure summary for by site
names(CN)

```{r}
std.error <- function(x) sqrt(var(x,na.rm=TRUE)/length(na.omit(x)))
soil.summary_site <- CN %>% group_by(n.treat, weevil, site)%>%
    summarise_at(c("d13c_21","C_21","d15n_21","d15n_21","N_21","C_18","N_18", "bd.pits.g.cm3.2018","sum.13","sum.14","sum.21","litter_kg_m2_mean","herb_kg_m2_mean","som_kg_m2_mean","litter_kg_m2_std.error" ,"herb_kg_m2_std.error" ,"som_kg_m2_std.error","numberwillow","numberother","totalnumber","willowstemdiam","otherstemdiam",          "totalstemdiam"), 
               funs(mean, std.error))
```
 
 Select only primary for site

```{r}
pri_cn_norep_site <- filter(soil.summary_site, site=="East"|site=="West")
write.csv(pri_cn_norep_site, "pri_cn_norep_site.csv")
pri_cn_norep_site <- read.csv("pri_cn_norep_site.csv")

pri_cn_norep_site$n.treat <- factor(pri_cn_norep_site$n.treat)


```

Select only secondary for site

```{r}
sec_cn_norep_site <- filter(soil.summary_site, site=="Norway"|site=="Harmony")
write.csv(sec_cn_norep_site, "sec_cn_norep_site.csv")
sec_cn_norep_site <- read.csv("sec_cn_norep_site.csv")
sec_cn_norep_site$n.treat <- factor(sec_cn_norep_site$n.treat)
```


#######################
##########################

STATS

primary and secondary 0 and 16 N only 
```{r}

soilc_21_site <- lmer(log(C_21)~n.treat+weevil+succession+weevil:succession+n.treat:succession+(1|plot.unq),data=CN_all)
anova(soilc_21_site)
summary(soilc_21_site)

step(soilc_21_site)

soilc_21_site <- lmer(log(C_21)~succession+weevil+(1|plot.unq),data=CN_all)
anova(soilc_21_site)
summary(soilc_21_site)

r.squaredGLMM(soilc_21_site)
plot(soilc_21_site)
qqnorm(resid(soilc_21_site))   
qqline(resid(soilc_21_site))

```

Primary Stats 

Carbon east west stats


SOIL C 2021
FINAL C MODEL
```{r}
names(pri_cn_rep)
soilc_21_site <- lmer(log(X21_mg.c.gsoil)~n.treat*weevil+(1|plot.unq),data=pri_cn_rep)
anova(soilc_21_site)
summary(soilc_21_site)

step(soilc_21_site)
soilc_21_site <- lmer(log(X21_mg.c.gsoil)~weevil+(1|plot.unq),data=pri_cn_rep)
anova(soilc_21_site)
summary(soilc_21_site)

r.squaredGLMM(soilc_21_site)
plot(soilc_21_site)
qqnorm(resid(soilc_21_site))   
qqline(resid(soilc_21_site))

```


SOIL N 2021
Final N pri sec 
```{r}
soiln_21_site <- lmer(log(N_21)~n.treat+weevil+succession+succession:n.treat+succession:weevil+(1|plot.unq),data=CN_all)
anova(soiln_21_site)
summary(soiln_21_site)


r.squaredGLMM(soiln_21_site)
plot(soilc_21_site)
qqnorm(resid(soilc_21_site))   
qqline(resid(soilc_21_site))
```


N east west stats


SOIL N 2021
FINAL N MODEL
```{r}
names(pri_cn_rep)
soiln_21_site <- lmer(log(X21_mg.n.gsoil)~n.treat*weevil+(1|plot.unq),data=pri_cn_rep)
anova(soiln_21_site)
summary(soiln_21_site)

r.squaredGLMM(soilc_21_site)
plot(soilc_21_site)
qqnorm(resid(soilc_21_site))   
qqline(resid(soilc_21_site))

```

Final d13c model pri sec 
```{r}
soil_d13c_21 <- lmer(d13c_21~n.treat+weevil+succession+weevil:succession+weevil:succession+(1|plot.unq),data=CN_all)
anova(soil_d13c_21)
summary(soil_d13c_21)
plot(soil_d13c_21)
qqnorm(resid(soil_d13c_21))  
qqline(resid(soil_d13c_21)) 
```

SOIL d13c 2021 primary
```{r}
names(pri_cn_rep)
str(pri_cn_rep)

soil_d13c_21<- lmer(X21_d13c~n.treat*weevil+(1|plot.unq),data=pri_cn_rep)
anova(soil_d13c_21)
summary(soil_d13c_21)
plot(soil_d13c_21)
qqnorm(resid(soil_d13c_21))   
qqline(resid(soil_d13c_21))
```



Final d15n model pri sec 
```{r}
soil_d15n_21 <- lmer(d15n_21~n.treat+weevil+succession+weevil:succession+weevil:succession+(1|plot.unq),data=CN_all)
anova(soil_d15n_21)
summary(soil_d15n_21)
plot(soil_d15n_21)
qqnorm(resid(soil_d15n_21))  
qqline(resid(soil_d15n_21)) 
```


SOIL d15n 2021 primary
```{r}
soil_d15n_21<- lmer(X21_d15n~n.treat*weevil+(1|plot.unq),data=pri_cn_rep)
anova(soil_d15n_21)
summary(soil_d15n_21)
plot(soil_d15n_21)
qqnorm(resid(soil_d15n_21))   
qqline(resid(soil_d15n_21))
```



##########################################
############################################
############################################

EFFECTS OF PH MOISUTRE BD ON SOIL C N D13C D15 ph, bd, moist effects on soil c 2021 across pri and sec NO EFFECT 
```{r}
C_21_ph_bd_moist_sec <- lmer(log(C_21)~pH2019+percent.moist2019+bd.cores.g.cm3.2021+(1|plot.unq),data=CN_all)
anova(C_21_ph_bd_moist_sec)
summary(C_21_ph_bd_moist_sec)
plot(C_21_ph_bd_moist_sec)
qqnorm(resid(C_21_ph_bd_moist_sec))  
qqline(resid(C_21_ph_bd_moist_sec))
```


ph, bd, moist effects on soil N 2021 NO EFFECT 
```{r}
N_21_ph_bd_moist_sec <- lmer(log(N_21)~pH2019+percent.moist2019+bd.cores.g.cm3.2021+(1|plot.unq),data=CN_all)
anova(N_21_ph_bd_moist_sec)
summary(N_21_ph_bd_moist_sec)
plot(N_21_ph_bd_moist_sec)
qqnorm(resid(N_21_ph_bd_moist_sec))  
qqline(resid(N_21_ph_bd_moist_sec))
```

ph, bd, moist effects on d13c 2021 NO EFFECT 
```{r}
d13c_21_ph_bd_moist_sec <- lmer(d13c_21~pH2019+percent.moist2019+bd.cores.g.cm3.2021+(1|plot.unq),data=CN_all)
anova(d13c_21_ph_bd_moist_sec)
summary(d13c_21_ph_bd_moist_sec)
plot(d13c_21_ph_bd_moist_sec)
qqnorm(resid(d13c_21_ph_bd_moist_sec))  
qqline(resid(d13c_21_ph_bd_moist_sec))
```

ph, bd, moist effects on d15n 2021 NO EFFECT
```{r}
d15n_21_ph_bd_moist_sec <- lmer(d15n_21~pH2019+percent.moist2019+bd.cores.g.cm3.2021+(1|plot.unq),data=CN_all)
anova(d15n_21_ph_bd_moist_sec)
summary(d15n_21_ph_bd_moist_sec)
plot(d15n_21_ph_bd_moist_sec)
qqnorm(resid(d15n_21_ph_bd_moist_sec))  
qqline(resid(d15n_21_ph_bd_moist_sec))
```


################################
################################
#############################
##############################################
LITTER AND HERB EFFECTS ON SOIL C,N, D13C, D135 PRIMARY DON'T HAVE THIS DATA FOR SECONDARY


###############################
#####################################
###################################
LITTER AND HERB STATS EFFECTS OF TREATMENTS 

data
names(CN)
```{r}
std.error <- function(x) sqrt(var(x,na.rm=TRUE)/length(na.omit(x)))
pitdata_2018 <- CN_all %>% group_by(n.treat, weevil, site, plot.unq, block)%>%
summarise_at(c("d13c_21","C_21","d15n_21","d15n_21","N_21","C_18","N_18", "sum.13","sum.14","sum.21","litter_kg_m2_mean","herb_kg_m2_mean","som_kg_m2_mean","litter_kg_m2_std.error" ,"herb_kg_m2_std.error" ,"som_kg_m2_std.error","pH2019", "percent.moist2019", "bd.cores.g.cm3.2021",    "bd.pits.g.cm3.2018"), 
               funs(mean, std.error))
               
pitdata_2018 <-pitdata_2018 %>% filter(site=="East"|site=="West")             
               
```
names(pitdata_2018)

LITTER MOSS MOD NO EFFECT 

```{r}
littermod_a<- lmer(log(litter_kg_m2_mean_mean)~n.treat*weevil+(1|block),data=pitdata_2018)
anova(littermod_a)
summary(littermod_a)
plot(littermod_a)
qqnorm(resid(littermod_a))  
qqline(resid(littermod_a))
```



LITTER MOSS MOD BY SITE NO EFFECT 
```{r}
littermod_b<- lmer(log(litter_kg_m2_mean_mean)~n.treat*weevil*site+(1|block),data=pitdata_2018)
anova(littermod_b)
summary(littermod_b)
plot(littermod_b)
qqnorm(resid(littermod_b))  
qqline(resid(littermod_b))
```


SOM MOD with site 
```{r}
#SOM
som_a<- lmer(log(som_kg_m2_mean_mean)~n.treat*weevil*site+(1|block),data=pitdata_2018)
anova(som_a)
summary(som_a)
plot(som_a)
qqnorm(resid(som_a))  
qqline(resid(som_a))
```

HERB MOD
```{r}
herb_a<- lmer(log(herb_kg_m2_mean_mean)~n.treat*weevil*site+(1|block),data=pitdata_2018)
anova(herb_a)
summary(herb_a)
plot(herb_a)
qqnorm(resid(herb_a))  
qqline(resid(herb_a))
```



###############################
####################################
#####################################

EFFECT of LITTER/MOSS, HERB, SOM  ON SOIL C AND N d13c d15n STATS only have this data for pri


Litter and herb and SOM and site effects on soil C 2021 NO EFFECT
```{r}
C_21_litter_herb_som_site <- lmer(log(C_21)~litter_kg_m2_mean+herb_kg_m2_mean+som_kg_m2_mean+som_kg_m2_mean:site+litter_kg_m2_mean:site+herb_kg_m2_mean:site+(1|plot.unq),data=pri_cn_rep)
anova(C_21_litter_herb_som_site)
summary(C_21_litter_herb_som_site)
plot(C_21_litter_herb_som_site)
qqnorm(resid(C_21_litter_herb_som_site))  
qqline(resid(C_21_litter_herb_som_site))


```

Litter and herb and SOM and site effects on soil N 2021 NO EFFECT
```{r}
N_21_litter_herb_som_site <- lmer(log(N_21)~litter_kg_m2_mean+herb_kg_m2_mean+som_kg_m2_mean+som_kg_m2_mean:site+litter_kg_m2_mean:site+herb_kg_m2_mean:site+(1|plot.unq),data=pri_cn_rep)
anova(N_21_litter_herb_som_site)
summary(N_21_litter_herb_som_site)
plot(N_21_litter_herb_som_site)
qqnorm(resid(N_21_litter_herb_som_site))  
qqline(resid(N_21_litter_herb_som_site))
```


litter and herb effects on soil by site d13c 2021 NO EFFECT 

```{r}
d13c_21_litter_herb_som_site <- lmer(d13c_21~litter_kg_m2_mean+herb_kg_m2_mean+som_kg_m2_mean+som_kg_m2_mean:site+litter_kg_m2_mean:site+herb_kg_m2_mean:site+(1|plot.unq),data=pri_cn_rep)
anova(d13c_21_litter_herb_som_site)
summary(d13c_21_litter_herb_som_site)
plot(d13c_21_litter_herb_som_site)
qqnorm(resid(d13c_21_litter_herb_som_site))  
qqline(resid(d13c_21_litter_herb_som_site))
```

litter and herb effects on soil by site d15n 2021 NO EFFECT 
```{r}
d15n_21_litter_herb_som_site <- lmer(d15n_21~litter_kg_m2_mean+herb_kg_m2_mean+som_kg_m2_mean+som_kg_m2_mean:site+litter_kg_m2_mean:site+herb_kg_m2_mean:site+(1|plot.unq),data=pri_cn_rep)
anova(d15n_21_litter_herb_som_site)
summary(d15n_21_litter_herb_som_site)
plot(d15n_21_litter_herb_som_site)
qqnorm(resid(d15n_21_litter_herb_som_site))  
qqline(resid(d15n_21_litter_herb_som_site))
```



##################################################
SOIL 2018 PITS EFFECTS OF TREATMENTS STATS

soil C 2018
```{r}
names(pri_cn_rep)
soilC_2018 <- lmer(log(X18_mg.c.gsoil)~n.treat*weevil+(1|plot.unq),data=pri_cn_rep)
anova(soilC_2018)
summary(soilC_2018)
plot(soilC_2018)
qqnorm(resid(soilC_2018))  
qqline(resid(soilC_2018))
```

soil N 2018
```{r}
soilN_2018 <- lmer(log(N_18)~n.treat*weevil*site+(1|plot.unq),data=pri_cn_rep)
anova(soilN_2018)
summary(soilN_2018)
plot(soilN_2018)
qqnorm(resid(soilN_2018))  
qqline(resid(soilN_2018))
```